\clearpage
\phantomsection
\addcontentsline{toc}{chapter}{ACKNOWLEDGEMENT}
\thispagestyle{footeronly}

\begin{center}
{\Large \textbf{ACKNOWLEDGEMENT}}
\end{center}

\justifying
\noindent

At the outset, we sincerely thank the Almighty for granting us the strength, wisdom, and perseverance to successfully complete this project.
We would like to convey our heartfelt gratitude to our respected project guide, Ms. Amrutha, Faculty of CUSAT, whose guidance, valuable suggestions, and continuous encouragement greatly influenced the progress and quality of our work. Her readiness to clarify our doubts and provide constructive feedback at every stage played a significant role in the successful completion of this project.
We also express our sincere gratitude to Dr. Pramod Pavithran, Head of the Department, for his constant support, encouragement, and for providing us with the opportunity and resources to carry out this project successfully.
We extend our sincere thanks to all the faculty members of CUSAT for their support and for providing the necessary facilities to carry out this work effectively.
Our sincere appreciation extends to our classmates and friends who supported us with their ideas, cooperation, and encouragement throughout the development of this project.
Finally, we express our deepest gratitude to our families for their unconditional support, patience, and belief in our abilities. Their encouragement has always been the foundation of our achievements.

\vspace{1.5cm}
\noindent
Mihal Palakkal\\
Ruvais P\\
Mohamed Jaseeb P J\\
Sajad Hussain Malla\\
Rishap Raj Chaudhary

\vfill
\clearpage