\chapter*{}
\thispagestyle{footeronly}
\addcontentsline{toc}{chapter}{ABSTRACT}
\begin{center}
{\Large \textbf{ABSTRACT}}
\end{center}

\noindent
This project proposes a Fully AI-Automated Washing Machine that intelligently analyzes clothes using a camera before initiating the washing process. The system captures images of the clothes placed inside the machine and applies computer vision and machine learning techniques to identify fabric type, color, dirt level, and load quantity. Based on this analysis, the system automatically determines the optimal amount of water, washing detergent, fragrance, wash cycle type, and wash duration required for effective cleaning. Unlike conventional washing machines that depend on manual user input or fixed wash programs, the proposed system dynamically adapts to each wash load, thereby minimizing water wastage, reducing detergent overuse, and lowering energy consumption. This approach improves washing efficiency while promoting environmental sustainability. The project aims to enhance user convenience by minimizing human intervention and ensuring consistent wash quality. Overall, the system demonstrates how artificial intelligence and automation can be integrated into household appliances to develop smarter, more efficient, and eco-friendly solutions.

\noindent
The system is implemented using a combination of hardware and software technologies. The hardware components include NodeMCU, relay modules, inlet and outlet diaphragm solenoid valves, a buzzer, and a stepper motor for controlling washing operations. The software stack consists of Flutter for the user interface, Python for image processing and decision logic, the Gemini-vision model, Firebase real-time database, and Arduino IDE.

\vspace{0.5cm}
\noindent
\textbf{Team Members:} \\
Mihal Palakkal \\
Ruvais P \\
Mohamed Jaseeb P J \\
Sajad Hussain Malla \\
Rishap Raj Chaudhary \\

\noindent
\textbf{Project Guide:} Ms. Amrutha

\vspace{0.5cm}
\noindent
\textbf{Keywords:} AI Washing Machine, Computer Vision, Fabric Recognition, Smart Home Appliances, Automated Wash Cycle, Water Conservation, Energy Efficiency, NodeMCU, Flutter, Firebase, Gemini-vision